Data splitting

### 2023-04-24 code/analysis_code/exploratory/exploratory_analysis.qmd
model building
fixing issues

```{r}
library(tidyverse)
library(here)
library(ggpmisc)
library(ggthemes)

@@

## Load data frames generated in data processing script

```{r}
here()
n1_n2_clean <- readRDS(here("data/processed_data/year2_cfx_clean.rds"))
plant_clean <- readRDS(here("data/processed_data/year2_plant_clean.rds"))
n1_n2_plant_combine <- readRDS(here("data/processed_data/year2_cfx_plant_combined.rds"))
covid_combine <- readRDS(here("data/processed_data/all_covid_combined.rds"))
all_data_combine <- readRDS(here("data/processed_data/year2_all_data_combined.rds"))
recovery_percent <- readRDS(here("data/processed_data/recovery_percentages.rds"))
n1_n2n2_avg <- readRDS(here("data/processed_data/year2_cfx_avg.rds"))
prop_pos_bio <- readRDS(here("data/processed_data/prop_pos_bio.rds"))
prop_pos_sample <- readRDS(here("data/processed_data/prop_pos_sample.rds"))

@@

### Look at `all_data_combine`

```{r}
tibble(all_data_combine)
```

@@

Graphs to make:

@@

Graphs to make:

## Single variate analysis

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos_7dma)) +
  geom_histogram(fill = "#C06C84") +
xlab(label = "Clinical Case Seven Day Moving Average") +
ylab(label = "Count") +
theme_light()

ggsave(here("figures/exploratory/variate/7DMA_histo.png"))
summary(covid_combine$pcr_pos_7dma)
```

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos)) +
  geom_histogram(fill = '#F8B195') +
  xlab(label = "Daily Positive Clinical Cases") +
  ylab(label = "Count") +
  theme_light()

ggsave(here("figures/exploratory/variate/pos_cases_histo.png"))
summary(covid_combine$pcr_pos)
```

### Clinical case 14 day moving average histogram

```{r}
covid_combine %>% ggplot(aes(pcr_pos_14dma)) +
  geom_histogram(fill = "#FF7E5A")

summary(covid_combine$pcr_pos_14dma)

@@

### Total copies in WW at county level histogram

```{r}
sum_daily_copies_county <- all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies))

  sum_daily_copies_county %>% ggplot(aes(log10(sum_total_copies), fill = target)) +
  geom_histogram() +
      scale_fill_manual(name = "Target",values = c("#F67280", "#355C7D")) +
    xlab(label = "Log10 Total Daily County-Level Copies") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/total_copies_histo.png"))
  log10_sum_daily_copies <- log10(sum_daily_copies_county$sum_total_copies)
  summary(log10_sum_daily_copies)
```

### Copies per liter histogram

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>%

  ggplot(aes(log10(sum_mean_copies_L), fill = target)) +
  geom_histogram() +
      scale_fill_manual(name = "Target",values = c("#F67280", "#355C7D")) +
    xlab(label = "Log10 Copies Per Liter") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/copies_L_histo.png"))

@@

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma),
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_7dma)) +
  geom_point(color = '#C06C84') +
  geom_line(color = '#C06C84') +
  xlab(label = "Date") +
  ylab(label = "Clinical Case 7-Day Moving Average") +
  theme_light()

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_pcr_pos)) +
  geom_bar(stat = "identity", color = '#F8B195') +
  geom_line(color = 'black') +
  ylab(label = "Positive Cases Per Day") +
  xlab(label = "Date") +
  theme_light()

@@

### Case 14 day moving average over time

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_14_dma = mean(pcr_pos_14dma)) %>%

   ggplot(aes(x = date, y = mean_14_dma)) +
  geom_bar(stat = "identity", color = '#FF7E5A') +
  geom_line() +
  ylab(label = "Clinical Case 14-Day Moving Average") +
  xlab(label = "Date") +
  theme_light()

@@

### Log10 viral load in wastewater over time (avg N1/N2)

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na(target) %>% #remove NAs from analysis
  group_by(date, facility) %>% #this will allow me to average the two targets
  summarize(mean_total_copies = mean(total_copies)) %>% #average across N1 and N2
  ungroup(facility) %>% #need to do this so I can add the avg total copies of each WRF
  summarize(sum_total_copies = sum(mean_total_copies)) %>% #total daily county copies
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point(color = "#80256D") +
  geom_line(color = "#80256D") +
   xlab(label = "Date") +
  ylab(label = "Log10 Total Daily County-Level Copies") +
  theme_light()

@@

### Log10 viral load in wastewater (both N1 and N2)

```{r}
all_data_combine %>% 
  group_by(date, target, facility) %>% 
  drop_na(target) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>%

  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  geom_line() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Date") +
  ylab(label = "Log10 County-Level Viral Load") +
  theme_light() +
  facet_wrap(vars(target),nrow = 2)

 ggsave(here("figures/exploratory/covariate/total_copies_time.png"))
```

### Log10 copies per liter across time (both N1 and N2)

```{r}
all_data_combine %>%
  drop_na(target) %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L),color = target)) +
geom_line() +
  geom_point() + 
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Date") +
  ylab(label = "Log10 County-Level Copies Per Liter") +
  theme_light() +
    facet_wrap(vars(target),nrow = 2)

 ggsave(here("figures/exploratory/covariate/copies_L_time.png"))

@@

```{r}
prop_pos_sample %>% group_by(date) %>% 
  summarize(avg_pos_rate = mean(pos_rate_tr)) %>%

  ggplot(aes(date, avg_pos_rate)) +
  geom_point() +
  geom_line()

@@

### Log10 copies per liter vs log10 case 7DMA (N1/N2 average)

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), log10(mean_7dma))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/copies_L_avg_vs_7DMA.png"))
```

### Log10 copies per liter vs log10 case 7DMA (N1 & N2)

```{r}
all_data_combine %>%  
  drop_na(target) %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), log10(mean_7dma), color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/copies_L_vs_7DMA.png"))
```

### Log10 Copies per Liter vs log10 daily clinical cases

```{r}
all_data_combine %>%  
  drop_na(target) %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies_L), log10(mean_pcr_pos), color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/copies_L_vs_pcr_pos.png"))
```

### Log10 Copies per Liter vs log10 daily clinical cases (N1 N2 average)

```{r}
all_data_combine %>%  
  drop_na(target) %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies_L), log10(mean_pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/copies_L_avg_vs_pcr_pos.png"))
```

### Log10 viral load vs log10 7DMA

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies), log10(mean_7dma))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 Daily Viral Load") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/avg_total_copies_vs_7DMA.png"))
```

### Log10 viral load vs log10 7DMA (N1 & N2)

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies), log10(mean_7dma), color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 Daily Viral Load") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_vs_7DMA.png"))
```

### Log10 viral load vs log10 daily positive cases

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies), log10(mean_pcr_pos), color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 Daily Viral Load") +
  ylab(label = "Log10 Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_vs_pcr_pos.png"))
```

### Log10 viral load vs log10 daily positive cases

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies), log10(mean_pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 Daily Viral Load") +
  ylab(label = "Log10 Daily Reported Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_avg_vs_pcr_pos.png"))
```

### Sample percent positive vs log10 7DMA

```{r}

>>>>

```{r, message=FALSE}
library(tidyverse)
library(here)
library(ggpmisc)
library(ggthemes)

@@

## Load final data frame generated in data processing script

```{r}
data_clean <- readRDS(here("data/processed_data/final_sum_combined_data.RDS"))

@@

### Look at `data_clean`

```{r}
tibble(data_clean)
```

@@

@@

-   Scatter plot of wastewater and clinical variables over time

-   Linear regressions between wastewater and clinical variables

## Variation analysis

@@

```{r}
data_clean %>% ggplot(aes((case_7dma))) +
  geom_histogram(fill = "#C06C84") +
xlab(label = "Clinical Case Seven Day Moving Average") +
ylab(label = "Count") +
theme_light()

ggsave(here("figures/exploratory/variate/7DMA_histo.png"))
summary(data_clean$case_7dma)
```

Slight bimodal distribution in 7DMA

@@

```{r}
data_clean %>% ggplot(aes(pcr_pos)) +
  geom_histogram(fill = '#F8B195') +
  xlab(label = "Daily Positive Clinical Cases") +
  ylab(label = "Count") +
  theme_light()

ggsave(here("figures/exploratory/variate/pos_cases_histo.png"))
summary(data_clean$pcr_pos)
```

The distribution of daily cases is heavily skewed to the right. Log-transforming gives a more normal distribution

### Clinical case percent positive tests

```{r}
data_clean %>% ggplot(aes(percent_pos_test)) +
  geom_histogram(fill = "#FF7E5A")

summary(data_clean$percent_pos_test)

@@

There is also a slight bimodal distribution with percent of positive tests, but nonetheless the distribution is right-skewed.

### Wastewater viral load histogram

```{r}
    data_clean %>% ggplot(aes(avg_VL)) +
  geom_histogram() +
    xlab(label = "Wastewater viral load") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/VL_histo.png"))

@@

Wastewater viral load is also very right-skewed. We may need to log-transform some variables before doing more statistical analysis.

### Log10 viral load histogram

```{r}
data_clean %>% 
  ggplot(aes(log10_VL)) +
  geom_histogram()

ggsave(here("figures/exploratory/variate/log10VL_histo.png"))

### Assay percent positivity histogram

```{r}
  data_clean %>% ggplot(aes(pos_rate_tr)) +
  geom_histogram() +
    xlab(label = "qPCR assay positivity") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/assay_pos_histo.png"))

Again with the bimodal distribution, even more so this time.

@@

```{r, message=FALSE}
data_clean %>% 
  ggplot(aes(x = date, y = case_7dma)) +
  geom_line(size =1.5, color = '#4c005c') +
  xlab(label = "Date") +
  ylab(label = "Clinical Case 7-Day Moving Average") +
  theme_light()

@@

```{r}
data_clean %>% 
  ggplot(aes(x = date, y = pcr_pos)) +
  geom_line(color = '#4c005c', size = 2) +
  ylab(label = "Positive Cases Per Day") +
  xlab(label = "Date") +
  theme_light()

@@

### Clinical case percent positive tests over time

```{r}
data_clean %>% 
   ggplot(aes(x = date, y = percent_pos_test)) +
  geom_line(size = 1.5) +
  ylab(label = "Percent of positive clinical tests") +
  xlab(label = "Date") +
  theme_light()

@@

### Wastewater viral load over time

```{r}
data_clean %>% 
  ggplot(aes(date, avg_VL)) +
  geom_line(color = "#2bce48", size = 1.5) +
   xlab(label = "Date") +
  ylab(label = "Total Daily County-Level Copies") +
  theme_light()

@@

### Log10 wastewater viral load over time

```{r}
data_clean %>% 
  ggplot(aes(date, log10_VL)) +
  geom_line(color = "#2bce48", size = 1.5) +
   xlab(label = "Date") +
  ylab(label = "Log10 Total Daily County-Level Copies") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_time.png"))
```

```{r, message=FALSE}
data_clean %>% 
  ggplot(aes(date, pos_rate_tr)) +
geom_line(size =1.5, color = '#634499') +
  xlab(label = "Date") +
  ylab(label = "Percent Positive N1 & N2 assays") +
  theme_light()

ggsave(here("figures/exploratory/covariate/percent_pos_time_plot.png"))

@@

### Tests administered over time

```{r}
data_clean %>% ggplot(aes(date, test_admin)) +
  geom_line(size = 1.5) +
  theme_light()

ggsave(here("figures/exploratory/covariate/test_admin_time.png"))
```

Tests administered drop off significantly after the beginning of 2022. In all other time series, there is a spike in mid-2022, but that spike is missing here.

From here on, I will split the data before and after Februrary 28, 2022 to see how correlations change.

@@

### Log10 wastewater viral load vs case 7DMA

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>% 
  ggplot(aes(log10_VL, case_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
   xlab(label = "Log10 County-Level Copies Per Liter") +
  ylab(label = "Log10 Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_vs_7DMA_Y2.png"))
```

RSQ value does not change between full data and the subset of dates

### Log10 wastewater viral load vs tests administered

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>% 
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>% 
  ggplot(aes(log10_VL, test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 viral load") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_testadmin_Y2.png"))
```

Correlations overall are weak, but the strongest positive correlation is found in the earlier dates

### Log10 wastewater viral load vs percent of positive tests

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(log10_VL, percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "Log10 wastewater viral load") +
  ylab(label = "Percent of positive clinical cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_percentcase_Y2.png"))
```

RSQ is slightly lower in later dates than in full data set, but does not change in earlier dates

### Log10 wastewater viral load vs daily reported cases

```{r}
#full dataset
data_clean %>%  
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases.png"))

#data split before FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases_Y1.png"))

#data split after FEB 28 2022
data_clean %>%  
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(log10_VL, log10(pcr_pos))) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq(color = "#80256D") +
  xlab(label = "Log10 wastwater viral load") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/log10VL_dailycases_Y2.png"))
```

RSQ is higher is the data splits than in the full set. I'm not sure what to think of this, but it could be due to the high error associated with this metric.

### Assay positivity vs 7DMA

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr, case_7dma)) +
  geom_point(color = "#00b0ba") +
  stat_poly_line(color = "#00b0ba") +
  stat_poly_eq(color = "black") +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_vs_7DMA_Y2.png"))
```

No significant change in data splits

### Assay positivity vs tests administered

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,test_admin)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR tests administered") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testadmin_Y2.png"))
```

RSQ is low across all data sets, but highest in earlier dates than in full data or later dates

### Assay positivity vs percent of positive tests

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,percent_pos_test)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "PCR test positivity") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_testpos_Y2.png"))
```

There is a slight decrease in RSQ in the later dates than in the full set or the earlier dates.

### Assay positivity vs daily reported cases

```{r}
#full dataset
data_clean %>% 
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases.png"))

#data split before FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) %>%
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases_Y1.png"))

#data split after FEB 28 2022
data_clean %>% 
  filter(between(date, as.Date("2022-02-28"), as.Date("2022-12-21"))) %>%
  ggplot(aes(pos_rate_tr,pcr_pos)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  xlab(label = "qPCR assay positivity") +
  ylab(label = "Daily reported cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/assaypos_dailycases_Y2.png"))
```

The highest RSQ value was in the earlier dates than in the full set or later dates

### 2023-04-24 code/modeling_code/1_data_splitting.Rmd
fixing issues

```{r}
library(tidyverse) 
library(tidymodels)
library(here)
library(skimr)
```

@@

```{r}
covid_ww_data <- readRDS(here("data/processed_data/final_sum_combined_data.RDS"))
#covid_ww_data <- covid_ww_data %>% drop_na()
tibble(covid_ww_data)
```

Since our data set is large enough, we can split it into a training set and a testing set.

First though, I want to split the data such that one set contains observations prior to July 2022 and the latter contains observations after July 2022. From my exploratory analysis, namely, comparing temporal trends of wastewater viral load with clinical case moving averages, I noted that there appears to be a disconnect in trends after the Omicron BA.5 wave (Late summer/fall 2022). As such, I would like to build models using data prior to this time and then test the model on the latter portion of the data to see if there is a difference.

@@

```{r}
set.seed(420)
train_dates <- covid_ww_data %>% filter(collection_num <= 193)

test_dates <- covid_ww_data %>% filter(collection_num > 193)

saveRDS(test_dates, here("data/modeling_data/test_dates.RDS"))
```

@@

```{r}
data_split <- initial_split(train_dates, prop = 3/4)

train_data <- training(data_split)
test_data <- testing(data_split)

>>>>

```{r, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(here)
library(skimr)
```

@@

```{r}
covid_ww_data <- readRDS(here("data/processed_data/final_sum_combined_data.RDS"))
covid_ww_data <- covid_ww_data %>% drop_na()
tibble(covid_ww_data)
```

I want to split the data such that one set contains observations prior to February 28, 2022 and the latter contains observations after February 28, 2022. From my exploratory analysis, namely, comparing temporal trends of wastewater viral load with clinical case moving averages, I noted that there appears to be a disconnect in trends after the Omicron BA.5 wave (Late summer/fall 2022). As such, I would like to build models using data prior to this time and then test the model on the latter portion of the data.

@@

```{r}
train_dates <- covid_ww_data %>% filter(between(date, as.Date("2020-06-30"), as.Date("2022-02-28"))) # n=151

test_dates <- covid_ww_data %>% filter(between(date, as.Date("2022-03-02"), as.Date("2022-12-21"))) # n=73

saveRDS(test_dates, here("data/modeling_data/test_dates.RDS"))

```

@@

```{r}
set.seed(420)
data_split <- initial_split(train_dates, prop = 3/4)

train_data <- training(data_split) # n=113
test_data <- testing(data_split) # n=38
